---
title: "PSTAT175 Project"
author: "Yuchen Zheng"
date: "11/24/2020"
output: word_document
---


```{r}
library(survminer)
library(ggplot2)
library(survival)
library(dplyr)
#convert numberic variables to factors 
colon$sex <- factor(colon$sex) 
colon$obstruct <- as.factor(colon$obstruct) 
colon$perfor<- factor(colon$perfor) 
colon$adhere <- factor(colon$adhere) 
colon$differ <- factor(colon$differ)
colon$extent <- factor(colon$extent) 
colon$surg <- factor(colon$surg) 
colon$node4 <- factor(colon$node4) 
colon$etype<- factor(colon$etype)
```

```{r}
#subset recurrence data
colon.recurrence <- subset(colon, etype == 1, select=c(id:etype))
```

```{r}
#plot KM estimate for recurrence data
recurrence.surv <- Surv(colon.recurrence$time, colon.recurrence$status)
recurrence.fit <- survfit(recurrence.surv~rx, data=colon.recurrence)
ggsurvplot(recurrence.fit, conf.int=F,
           title = 'Kaplan-Meier Curve for Colon Cancer \nRecurrence by Treatment',
           xlab= 'Times (until death) \n in Days')
```
```{r}
r.rx.coxph <- coxph(Surv(time, status) ~ rx, data=colon.recurrence)
summary(r.rx.coxph)
```
```{r}
#removing NA values
colon.recurrence1 = na.omit(colon.recurrence)
```

```{r}
r.model1 <- coxph(Surv(time, status) ~ sex, data=colon.recurrence1)
r.model2 <- coxph(Surv(time, status) ~ age, data=colon.recurrence1)
r.model3 <- coxph(Surv(time, status) ~ obstruct, data=colon.recurrence1)
r.model4 <- coxph(Surv(time, status) ~ perfor, data=colon.recurrence1)
r.model5 <- coxph(Surv(time, status) ~ adhere, data=colon.recurrence1)
r.model6 <- coxph(Surv(time, status) ~ differ, data=colon.recurrence1)
r.model7 <- coxph(Surv(time, status) ~ extent, data=colon.recurrence1)
r.model8 <- coxph(Surv(time, status) ~ surg, data=colon.recurrence1)
r.model9 <- coxph(Surv(time, status) ~ node4, data=colon.recurrence1)
BIC(r.model1, r.model2,r.model3, r.model4,r.model5,r.model6,r.model7,r.model8,r.model9)
```

```{r}
r.model9.1 <- coxph(Surv(time, status) ~ node4+sex, data=colon.recurrence1)
r.model9.2 <- coxph(Surv(time, status) ~ node4+age, data=colon.recurrence1)
r.model9.3 <- coxph(Surv(time, status) ~ node4+obstruct, data=colon.recurrence1)
r.model9.4 <- coxph(Surv(time, status) ~ node4+perfor, data=colon.recurrence1)
r.model9.5 <- coxph(Surv(time, status) ~ node4+adhere, data=colon.recurrence1)
r.model9.6 <- coxph(Surv(time, status) ~ node4+differ, data=colon.recurrence1)
r.model9.7 <- coxph(Surv(time, status) ~ node4+extent, data=colon.recurrence1)
r.model9.8 <- coxph(Surv(time, status) ~ node4+surg, data=colon.recurrence1)
BIC(r.model9.1,r.model9.2,r.model9.3,r.model9.4,r.model9.5,r.model9.6,r.model9.7,r.model9.8)

```

```{r}
r.model9.7.1 <- coxph(Surv(time, status) ~ node4+extent+sex, data=colon.recurrence1)
r.model9.7.2 <- coxph(Surv(time, status) ~ node4+extent+age, data=colon.recurrence1)
r.model9.7.3 <- coxph(Surv(time, status) ~ node4+extent+obstruct, data=colon.recurrence1)
r.model9.7.4 <- coxph(Surv(time, status) ~ node4+extent+perfor, data=colon.recurrence1)
r.model9.7.5 <- coxph(Surv(time, status) ~ node4+extent+adhere, data=colon.recurrence1)
r.model9.7.6 <- coxph(Surv(time, status) ~ node4+extent+differ, data=colon.recurrence1)
r.model9.7.7 <- coxph(Surv(time, status) ~ node4+extent+surg, data=colon.recurrence1)
BIC(r.model9.7.1,r.model9.7.2,r.model9.7.3, r.model9.7.4,r.model9.7.5,r.model9.7.6,r.model9.7.7)
```


```{r}
r.model9.7.7.1 <- coxph(Surv(time, status) ~ node4+extent+surg+sex, data=colon.recurrence1)
r.model9.7.7.2 <- coxph(Surv(time, status) ~ node4+extent+surg+age, data=colon.recurrence1)
r.model9.7.7.3 <- coxph(Surv(time, status) ~ node4+extent+surg+obstruct, data=colon.recurrence1)
r.model9.7.7.4 <- coxph(Surv(time, status) ~ node4+extent+surg+perfor, data=colon.recurrence1)
r.model9.7.7.5 <- coxph(Surv(time, status) ~ node4+extent+surg+adhere, data=colon.recurrence1)
r.model9.7.7.6 <- coxph(Surv(time, status) ~ node4+extent+surg+differ, data=colon.recurrence1)
BIC(r.model9.7.7.1, r.model9.7.7.2,r.model9.7.7.3,r.model9.7.7.4,r.model9.7.7.5,r.model9.7.7.6)
```


```{r}
#full model
r.model.full <- coxph(Surv(time, status) ~ sex + age + obstruct + perfor + adhere + differ + extent + surg + node4, data=colon.recurrence1)
```

```{r}
BIC(r.model.full,r.model9.7.7.3,r.model9.7.7,r.model9.7,r.model9)
```
We can use the Analysis of Deviance procedure to get the proper Likelihood Ratio Test
```{r}
anova(r.model9.7.7)
```
```{r}
r.model9.7.7
```

```{r}
colon.recurrence2 <- colon.recurrence
colon.recurrence2$nodes[is.na(colon.recurrence2$nodes)] <- mean(colon.recurrence2$nodes, na.rm = TRUE)
colon.recurrence2$differ[is.na(colon.recurrence2$differ)] <- 2
colon.recurrence2$differ <- factor(colon.recurrence2$differ, exclude=NULL)
```
We created to a second recurrence dataset by replcing the na values in nodes with the mean of the column and include the NA value as a character level in the differ column to check the cox proportional model assumption.

```{r}
recurrence2.surv <- Surv(colon.recurrence2$time, colon.recurrence2$status)
plot(survfit(recurrence2.surv~rx, data=colon.recurrence2),
     fun='cloglog',
     col =c(2,3,4),
     ylab='log-log(S(t))',
     xlab='Time(days)',
     main="C-log-log plot for Covariate rx")
legend("topleft" ,legend =c("Obs","Lev","Lev+5FU"), col=c(2,3,4),lty=1)
```
The three curves in the cloglog plot for covariate rx is overlapping before 100 days, but after 100 days the two curves for Obs and Lev are crossing over and the two curves appear to be parallel to the curve for Lev+5Fu. We are somewhat concerning that the assumption might be violated because of the crossover before 100 days. 

```{r}
plot(survfit(recurrence2.surv~surg, data=colon.recurrence2),
     fun='cloglog',
     col =c(2,3),
     ylab='log-log(S(t))',
     xlab='Time(days)',
     main="C-log-log plot for Covariate surg")
legend("topleft" ,legend =c("short","long"), col=c(2,3),lty=1)
```
The two curves in the C-log-log plot for covariate surg is crossing over and very close to each other. We conclude that the assumption is approriate to use for this covariate because crossingover means the two groups have the same hazard ratio. 
```{r}
plot(survfit(recurrence2.surv~extent, data=colon.recurrence2),
     fun='cloglog',
     col =c(2,3,4,5),
     ylab='log-log(S(t))',
     xlab='Time(days)',
     main="C-log-log plot for Covariate extent")
legend("topleft" ,legend =c("1-submucosa","2-muscle","3-serosa", "4-contiguous structures"), col=c(2,3,4,5),lty=1)
```
The curves in the C-log-log plot are crossing over after 100 days. Since there are not enough data points in each group to show a more comprehesive trend, it's hard for us to make a decision based on the plot. 
```{r}
plot(survfit(recurrence2.surv~node4, data=colon.recurrence2),
     fun='cloglog',
     col =c(2,3),
     ylab='log-log(S(t))',
     xlab='Time(days)',
     main="C-log-log plot for Covariate node4")
legend("topleft" ,legend =c("more than 4 postivie nodes","less than 4 nodes"), col=c(2,3),lty=1)
```
The two curves in this C-log-log plot appear to be parallel to each other, so we think the cox proportional assumption is approritate for the covariate node4. 
```{r}
cox.zph(coxph(formula = Surv(time, status) ~ node4 +extent+surg+rx, data = colon.recurrence2))
```
Since the p value for node4 is less than 0.05, there is significant evidence that the cox proportional model assumption is violated for variable node4. However, the C-log-log plot for covariate node4 looks parallel, so we think the significance of p-value might be due to some noise in the data at the beginning of the study. Additionally, the overall p-value is greater than 0.05 which means we fail to reject the null hypothesis and can conclude that the cox proportional model assumption is reasonable to use for this model. 



```{r}
colon.recurrence.coxph <- coxph(Surv(time, status) ~ node4+extent++surg+rx, data=colon.recurrence1)
summary(colon.recurrence.coxph)
```
```{r}
anova(colon.recurrence.coxph)
```








#Marginal model for Death
```{r}
colon.death <- subset(colon, etype == 2)

death.fit <- survfit(Surv(time,status) ~ rx, data = colon.death)
ggsurvplot(death.fit, conf.int = F, 
           title = "Kaplan-Meier Curve for Colon Cancer Mortality \nby Treatment", 
           xlab = "Time (until death) \n in Days")
```

```{r}
d.rx.coxph <- coxph(Surv(time, status) ~ rx, data=colon.death)
summary(d.rx.coxph)
```

```{r}
#removing NA values
colon.death1 <- na.omit(colon.death)
```

```{r}
d.model1 <- coxph(Surv(time, status) ~ sex, data=colon.death1)
d.model2 <- coxph(Surv(time, status) ~ age, data=colon.death1)
d.model3 <- coxph(Surv(time, status) ~ obstruct, data=colon.death1)
d.model4 <- coxph(Surv(time, status) ~ perfor, data=colon.death1)
d.model5 <- coxph(Surv(time, status) ~ adhere, data=colon.death1)
d.model6 <- coxph(Surv(time, status) ~ node4, data=colon.death1)
d.model7 <- coxph(Surv(time, status) ~ differ, data=colon.death1)
d.model8 <- coxph(Surv(time, status) ~ extent, data=colon.death1)
d.model9 <- coxph(Surv(time, status) ~ surg, data=colon.death1)


BIC(d.model1, d.model2, d.model3, d.model4, d.model5, d.model6, d.model7, d.model8, d.model9)
```

```{r} 
d.model6.1 <- coxph(Surv(time, status) ~ node4 + sex, data=colon.death1)
d.model6.2 <- coxph(Surv(time, status) ~ node4 + age, data=colon.death1)
d.model6.3 <- coxph(Surv(time, status) ~ node4 + obstruct, data=colon.death1)
d.model6.4 <- coxph(Surv(time, status) ~ node4 + perfor, data=colon.death1)
d.model6.5 <- coxph(Surv(time, status) ~ node4 + adhere, data=colon.death1)
d.model6.6 <- coxph(Surv(time, status) ~ node4 + surg, data=colon.death1)
d.model6.7 <- coxph(Surv(time, status) ~ node4 + differ, data=colon.death1)
d.model6.8 <- coxph(Surv(time, status) ~ node4 + extent, data=colon.death1)

BIC(d.model6.1, d.model6.2, d.model6.3, d.model6.4, d.model6.5, d.model6.6, d.model6.7, d.model6.8)
```


```{r}
d.model6.8.1 <- coxph(Surv(time, status) ~ node4 + extent + sex, data=colon.death1)
d.model6.8.2 <- coxph(Surv(time, status) ~ node4 + extent + age, data=colon.death1)
d.model6.8.3 <- coxph(Surv(time, status) ~ node4 + extent + obstruct, data=colon.death1)
d.model6.8.4 <- coxph(Surv(time, status) ~ node4 + extent + perfor, data=colon.death1)
d.model6.8.5 <- coxph(Surv(time, status) ~ node4 + extent + adhere, data=colon.death1)
d.model6.8.6 <- coxph(Surv(time, status) ~ node4 + extent + differ, data=colon.death1)
d.model6.8.7 <- coxph(Surv(time, status) ~ node4 + extent + surg, data=colon.death1)

BIC(d.model6.8.1, d.model6.8.2, d.model6.8.3, d.model6.8.4, d.model6.8.5, d.model6.8.6, d.model6.8.7)
```



```{r}
d.model.full <- coxph(Surv(time, status) ~ sex + age + obstruct + perfor + adhere + differ + extent + surg + node4, data=colon.death1)
```

```{r}
BIC(d.model.full,d.model6.8.7, d.model6.8, d.model6)
```

```{r}
anova(d.model6.8)
```

```{r}
colon.death2 <- colon.death
colon.death2$nodes[is.na(colon.death2$nodes)] <- mean(colon.death2$nodes, na.rm = TRUE)

colon.death2$differ <- factor(colon.death2$differ, exclude=NULL)
```




```{r}
dnode4.fit <- survfit(Surv(time, status) ~ node4, data = colon.death2)

ggsurvplot(dnode4.fit , conf.int = F, 
           fun = "cloglog",
           xlim = c(20, 5000),
           title = "C-Log-Log for Colon Cancer Mortality \nby node4", 
           xlab = "Time (until death) \n in Days")
```

```{r}
dextent.fit <- survfit(Surv(time, status) ~ extent, data = colon.death2)

ggsurvplot(dextent.fit , conf.int = F, 
           fun = "cloglog",
           xlim = c(20, 5000),
           title = "C-Log-Log for Colon Cancer Mortality \nby extent", 
           xlab = "Time (until death) \n in Days")
```

```{r}
dextent.fit <- survfit(Surv(time, status) ~ rx, data = colon.death2)

ggsurvplot(dextent.fit , conf.int = F, 
           fun = "cloglog",
           xlim = c(20, 5000),
           title = "C-Log-Log for Colon Cancer Mortality \nby rx", 
           xlab = "Time (until death) \n in Days")
```


```{r}
cox.zph(coxph(formula = Surv(time, status) ~ node4 +extent+rx, data = colon.death2))
```

```{r}
d.final.model <- coxph(Surv(time, status) ~ node4 + extent + rx, data=colon.death2)
summary(d.final.model)
```
```{r}
anova(d.final.model)
```
